\section{Background and Related work}
At the heart of Tisane is the goal to derive statistical models from conceptual
models. To do so, Tisane relies on transforming aspects of analysts' expressed
conceptual models into causal graphs. There are multiple frameworks for
reasoning about causality~\cite{rubin2004teaching,pearl1995causal}. One
widespread approach is to use directed acyclic graphs (DAGs) to encode
conditional dependencies between
variables~\cite{pearl1995doCalculus,greenland1999causal,spirtes1994conditional,spirtes1996using}.
If analysts can specify a formal causal graph, Pearl's ``backdoor path
criterion''~\cite{pearl1995causal,pearl2000models} explains the set of variables
that control for confounding. However, in practice, specifying proper causal
DAGs is challenging and error-prone for domain experts who are not also experts
in causal analysis~\cite{suzuki2020causal}. Empirical findings may be
inconclusive or ambiguous in the causal relationships they
suggest~\cite{suzuki2018mechanisms}. Statistical non-experts also lack guidance
on which variables and relationships to include~\cite{velentgas2013developing}.
Despite having important domain knowledge, analysts do not have interfaces that
allow them to express what they know in a way that is approachable to them.
Therefore, Tisane does not expect analysts to specify a formal causal graph.
Instead, analysts can express causal relationships as well as more ambiguous
relationships between variables in the \SDSLlong.
% Both kinds of relationships are necessary 

Furthermore, prior work in the causal reasoning literature shows how linear
models can be derived from causal graphs to make statistical inferences and test
the motivating causal graph~\cite{spirtes1996using,spirtes1994conditional}.
Recently, VanderWeele proposed the ``modified disjunctive cause
criterion''~\cite{vanderweele2019modifiedDisjunctiveCriterion} as a new
heuristic for researchers without a clearly accepted formal causal model to
identify confounders to include in a linear model, for example. The criterion
identifies confounders in a graph based on expressed causal relationships. The
first release of Tisane (\autoref{sec:tisane}) applies the modified disjunctive
cause criterion when suggesting variables to include in a GLM or GLMM. Tisane
does not automatically include variables to the statistical models because
substantive domain knowledge is necessary to resolve issues of temporal
dependence between variables, among other
considerations~\cite{vanderweele2019modifiedDisjunctiveCriterion}. In the second
release of \tisane (\autoref{sec:rTisane}), we use the more recent
recommendations from Cinelli, Forney, and Pearl~\cite{cinelli2020controls} for
controls in regression models. To guide analysts through the suggestions, Tisane
provides analysts with explanations to aid their decision making during
disambiguation. 

Importantly, generalized linear models with or without mixed effects are not
formal causal analyses. Tisane does not calculate average causal effect or other
causal estimands. Rather, Tisane only utilizes insights about the connection
between causal DAGs and linear models to guide analysts towards including
potentially relevant confounders in their GLMs grounded in domain knowledge. 


\subsection{Statistical scope}  \label{sec:GLM}
% \section{Why is Tisane necessary, isn't Tea enough?}
% Tisane supports two classes of models that are widely applicable to diverse
% domains and data collection
% settings~\cite{lo2015transform,barr2013random,bolker2009generalized}:
% Generalized Linear Models (GLMs) and Generalized Linear Mixed-effects Models
% (GLMMs).
% Many research questions analysts want to answer require more complex analyses.

Generalized linear models (GLMs) and generalized linear models with mixed
effects (GLMMs) are meaningful targets because they are commonly used (e.g., in
psychology~\cite{lo2015transform,cohen2013applied}, social
science~\cite{kreft1998introducing}, and
medicine~\cite{bolker2009generalized,barr2013random}) yet are easy to misspecify
for statistical experts and non-experts alike~\cite{barr2013random,
cohen2013applied}. We designed Tisane to support researchers who are domain
experts capable of supplying conceptual and data collection information but lack
the statistical expertise or confidence to author GLM/GLMMs accurately.
%capable of specifying accurate conceptual and
%data relationships.
Both GLMs and GLMMs consist of (i) a \textit{model effects structure},
which can include main and interaction effects and (ii) \textit{family} and
\textit{link} functions. The family function describes how the residuals of a
model are distributed. The link function transforms the predicted values of the
dependent variable. This allows modeling of linear and non-linear relationships
between the dependent variable and the predictors. In contrast to
transformations applied directly to the dependent variable, a link function does
not affect the error distributions around the predicted values. The key
difference between GLMs and GLMMs is that GLMMs contain random effects in their
model effects structure. Random effects describe how individuals (e.g., a study
participant) vary and are necessary in the presence of hierarchies, repeated
measures, and non-nesting composition
(\ref{sec:data-measurement-relationships})\footnote{Traditionally, the term
``mixed effects'' refers to the simultaneous presence of ``fixed'' and
``random'' effects in a single model. We try to avoid these terms as there are
many contradictory usages and definitions~\cite{gelmanFixedRandom}. When we do
use these terms, we use the definitions from Kreft and De
Leeuw~\cite{kreft1998introducing}.}.

Both GLMs and GLMMs assume that (i) the variables involved are linearly related,
(ii) there are no extreme outliers, and (iii) the family and link functions are
correctly specified. In addition, GLMs also assume that (iv) the observations
are independent. Tisane's interactive compilation process guides users through
specifying model effects structures, family and link functions to satisfy
assumption (iii), and random effects only when necessary to pick between GLMs
and GLMMs and satisfy assumption (iv).

We scoped Tisane to GLM and GLMMs because they encompass a large scope of
statistical models such that our research contributions are widely applicable
and substantial. In addition, given that GLMs and GLMMs can represent common
Null Hypothesis Significance Tests (in \tea), \tisane generalizes our approach
in \tea. \tisane gives further evidence of the benefit of conceptual programming
abstractions and automated reasoning for authoring statistical analyses.

% We output models, not causal measures/estimates because people might want to revise, iterate on. They also expect a model, so Tisane is just a step towards moving people towards more sophisticaed estimates? 

\begin{comment}
Recent work in the database community helps researchers answer causal questions
about multilevel, or hierarchical, data~\cite{salimi2020causal, kayali2020demonstration}.
% \footnote{Generalized linear models with mixed effects are appropriate in these settings as well.},
CaRL~\cite{salimi2020causal} provides a domain-specific language to express
causal relationships between variables and a GUI to show researchers %the
results. \tea and \tisane leverage a similar insight that researchers have domain
knowledge that a system can use to infer statistical methods. Whereas CaRL is
focused on answering specific queries about average causal effect, the systems
in this dissertation are designed to address a range of non-causal questions as well.
\end{comment}