% \begin{localsize}{10}
% \small
I am lucky to have many, many people to thank. 

\textit{I thank my advisors Jeffrey Heer and \reneJust.} Jeff, thank you for
your endless creativity, sense of humor, and focused attention in all our
conversations. Thanks also for helping me navigate personal and professional
challenges with integrity. I had a lot of fun and freedom during the second half of my PhD
thanks to you. \rene, thank you for your consistent insistence on
the details.

% \item Committee 
\textit{I have the best committee, hands down.} Emery Berger, I am deeply grateful
that you agreed to meet with me, then a total stranger, at the start of your
sabbatical at MSR in 2018. Who knew that coffee would lead to many, many
more espressos while we worked on Tea, Scone, and a number of other projects. You have been my most ardent
champion and creative conversational partner. Thank you for including me in on your latest crazy idea and laughing with me
through the highs and lows of graduate school. 
% Thank you for always keeping
% me updated on your latest crazy idea and bringing me into projects. Your
% support through the highs and lows of graduate school, always with a heavy
% bout of laughter, kept me going. 
% Emery, you may be the best mad scientist I know, but above all,
% you are the strongest advocate I could have asked for. 
Leilani Battle, thank you for your enthusiasm and support since you were a
postdoc at UW! It's great to have you back.    
Tyler McCormick, thank you for welcoming me into your research group and the
Center for Statistics and the Social Sciences (CSSS). Our conversations and
collaboration have expanded my thinking and scholarship. 
% \item IDL, PLSE, Zach 

\textit{I am grateful to have been part of the warm UW CSE community throughout
my PhD.} In particular, I am indebted to the members of the Interactive Data Lab
(IDL) with whom I overlapped during my PhD for their camaraderie: Arvind
Satyanarayan, Ham Wongsuphasawat, Dominik Moritz, Yeaseul Kim, Younghoon Kim,
Jane Hoffswell, Sherry Wu, Alex Kale, Yang Liu, Matt Conlen, Zening Qu, Mick
Kittivorawong, Ameya Patil, Junran Yang, Madeleine Grunde-McLaughlin, Josh
Horowitz, and Luke Snyder. Michael Correll and Maureen Stone, while not students
in IDL, were consistent members of the lab. Thank you, all, for keeping me
humble, honest, and hopeful about the positive social impact of great research
and open-source software.

% In addition to IDL, 
I have also been fortunate to be part of the idiosyncratic PLSE
lab. 
% To my friends in PLSE over the years, of which there are too many to name
% individually, 
% Thank you for all the snacks, heated
% conversations, and tea. 
To my PLSE labmates over the years, of which there are too many to name
individually, I love how ferociously curious and whimsical you are about
everything in research and life. I would not have pursued this
research without PLSE welcoming me early on. In this regard, I want to give
special thanks to my friends Sarah Chasins, James Bornholt, Mangpo Phothilimthana, Sami Davies, Chandra
Nandi, Amanda Swearngin, Chenglong Wang, Jasper Tran O'Leary, Gus Smith, Pavel Panchekha,
and Max Willsey. Zachary Tatlock, thank you for investing so much of yourself
into making PLSE the best place to pursue daring research, have fun, and make
friends. 

To Elise Dorough, thank you for everything you do to keep CSE running while
still making time to help my peers and me through any situation (and many
emotions). To Joe Eckert, thank you for answering my numerous Slack messages
and cries for help even when I caused my own logistical problems. To
Elle Brown, thank you for your kindness. To Lisa Merlin, thank you for
helping me with countless reimbursements and doing everything in your power
to protect me from bureaucratic processes.

% To my CSE 492 (?) officemates, especially James Bornholt and Callie Bee, thank
% you for sharing in many humorous memories. 
\textit{The larger UW ecosystem has also shaped my scholarship.} To the members
of CSSS, thank you for welcoming me and giving me many opportunities to learn
from and ask questions of scholars across the social and statistical sciences.
The interdisciplinary seminars and courses, such a Thomas Richardson's offering
of CSSS/STAT 566 on causal reasoning, were instrumental in my graduate
education. Thanks also to the larger HCI community in DUB. There are a few
individual mentors to whom my gratitude is overdue: Daniela Rosner (HCDE), Tim
Althoff (CSE), and Amy Ko (iSchool). To Daniela, especially, thank you for
sharing your brilliance, wisdom, and curiosity about ``design'' and life over
memorable chats in your office and on walks throughout the city. 
% I have been incredibly lucky to have mentors across
% campus. To Daniela Rosner, whose encouragement and mentorship during my PhD
% is long overdue, thank you for sharing your curiosity, creativity, and
% humor. To Tim Althoff, thank you for our brainstorming sessions to explore
% research ideas. 

% \item MSR internship mentors: Mary, Daniel, Ben 
\textit{During my PhD, I had so much fun and expanded my research horizons
through two summer internships at Microsoft Research.} Mary Czerwinski, Daniel
McDuff, and Ben Zorn, thank you for entrusting me to run with new ideas,
modeling constructive mentorship (which I strive to emulate), and becoming
friends over the years. 

% \item Mentees 
% \item IHME collaborators, CSSS 
\textit{To my collaborators, thank you for your inspiration, energy, and support.}
Specifically, I thank Joseph Dieleman and Goli Tsakalos, for inviting me as
a volunteer with the Institute for Health Metrics and Evaluation (IHME) the
last two years of my PhD. To Sawyer Crosby and Emily Johnson, thank you for
your time and energy in showing me the ropes at IHME, sharing your analysis
experiences, and trying out and giving helpful feedback on early prototypes
of my systems. To Maureen Daum, thank you for your candor, reliability, and
sense of humor as a collaborator and friend. To Nicole de Moura, Grace Oh,
Melissa Birchfield, Pranav Rajan, Shreyash Nigam, Josh Pollock, Annie
Denton, Blue A. Jo, Reiden Chea, Corinne Herzog, Vincent Pun, Irene Luo, Ken
Gu, Audrey Seo, and Edward Misback, thanks for entrusting me to mentor you during your
explorations of research and computer science. I am so immensely proud of
how you have pursued your interests, and I look forward to seeing
you continue to grow!

\textit{I would not have pursued a PhD without formative mentorship and research
experiences as an undergraduate}. I thank my Vanderbilt undergraduate advisors Bobby
Bodenheimer and John Rieser for encouraging me to embody and live out my
curiosities. To Richard Paris and Lauren Buck, my cool
graduate student labmates, thank you for your friendship and our memorable
dinner parties. I also thank the Ingram Scholarship Program for fostering my
commitment to civic engagement, practice to reflect and think critically, and
friendships that will continue to endure. To my REU mentors at the
University of Utah, Bill Thompson, Sarah Creem-Regehr, and Jeanine Stefanucci,
thank you for inviting me into your lab, sharing about life and research with
candor, and modeling collaboration and leadership. To Shashank Pandey and Jubi
Taneja, thank you for sharing poetry, meals, and laughter as we became friends
that formative summer at Utah.

% \item UW CSE + DUB, Daniela, Tim; officemates + batchmates; Elise, Joe, Elle, Lisa; Maureen (co-author)
% \item Council of Wise Women: Advent Anglican, St. Mark's gardening club, Seattle Mosaic Arts
\textit{I am immensely grateful to my community in and around Seattle.} To my
Council of Wise Women from Advent Anglican (especially Kristina Bresnen),
Seattle Mosaic Arts (I hear you laughing, June Chang and Claire Barnett!), and
St. Mark's Gardening Ministry (especially Kathy Sodergren, Kelly Lundquist,
and Kathy Thomason), thank you for enriching my soul. You have shared so much wisdom
in the countless hours I have sat in your presence, often in silence.
% even if there were no words exchanged. 
% , sometimes
% praying, piecing together mosaics, or pulling weeds (what is the
% difference?)

% \item The Emily Program; Patrice; Carrie and Traig and DBT group at Hall Health 
% I give my heartiest, warmest thank you to ``The Emotional B*tches
My heart is beyond a five on the fullness scale, overflowing with marbles for the
``The Emotional B*tches (Thoughts-Emotions-Behaviors)'' and the healthcare
professionals at The Emily Program. 
% You know who you are. 
% who made me laugh, rode
% the elevators up and down one floor with me, and created a true fellowship.
% Our fervor for life (and death) was palpable in everything. 
% beyond full full, overflowing with marbles. 
% My heart is beyond a 5 on the fullness scale. It's overflowing with marbles.
% I can still picture the group room on the eighth floor and incredible view
% of the duck dash on Lake Union every Tuesday night. 
Thank you for creating a true fellowship and, quite literally, keeping me alive.
Patrice Staiger, Carrie DeMartini, and Bethany Bylsma, thank you for guiding me
to, through, and from The Emily Program. Treg Isaacson, Hui Sun, and the others 
% I encountered 
at Hall Health, thank you for lighting the way.
% thank you for being points of light guiding my
% path.
% brightening my path
% always. 

% \item Friends, Besties 
\textit{I thank my friends and family.} To Bridget Claborn, thank you for your
incisive questions, belief in the poetry within and all around us (i.e.,
post-post-modernism, dahlias, anemones, Mary Oliver), and commitment to weave
together research and practice for social change, for life.
%to feel and live in the poetry of the universe. 
To Aoife Blacklaws, thank you for encouraging me in all my endeavors and sharing
your awe of every little creature. To Michael Zuch, thank you for sharing in the
tender, unanswerable questions of faith, theology, and the meaning of life. To Rachael
Grenfell-Dexter, thank you for sharing many adventures around the world and giving me permission to have fun, grow, and evolve. 
To Lucy Rahner, thank you for sharing your artistry and earnest search for the truth and
goodness in everything. To Catherine Chung, thank you for keeping it real,
talking with me endlessly for days, and listening so whole-heartedly. 
To Maria Ventura, thank you for being quick to laugh, judging a person's character based on their pasta, and inspiring me in research.
To Morelle Arian, thank you for welcoming me into your family and sharing your delight in wherever life takes us. 
To Olivia Bradley-Skill, thank you for encouraging me to lean into my creativity and explore ways of combining my interests ever since we were kids.
To the friends named and unnamed, I am honored to share this life with you. 
%incredible earth, even for a brief moment in time. %--how impeccable. 

% \item Partner, Family 
% and fight for
% Finally, to Jared Roesch, thank you for your love, commitment to community,
% and ambition to focus on what truly matters. 
Finally, to Jared Roesch, thank you for your love and support throughout the PhD.
To my parents Ok Za Julie and Stephen Jun,
thank you for your sacrifices and support. I love that you
encourage me to write better, more papers even when, by your estimation,
no one reads them.
% \end{itemize}
% \end{localsize}

% Advisors
% Emery

% CSSS, Tyler

% Committee

% I've been lucky to have the counsel of wise women in Seattle. 
% Kristina, Alison, Deborah
% St Mark's - gardening group
% Seattle Mosaic Arts

% The Emily Program - thank you for keeping me alive. 
% Use initials to refer to specific people
% I swear, our daily meals are amazing sources of inspiration for comedy.
% Cursing in my internal monologue

% IDL 

% PLSE

% Mentors - Zach, Daniela Rosner

% MSR - Mary, Daniel, Ben

% IHME 

% Mentees
% ....

% Larger research community
% CSE community - Elise, Joe, Lisa Merlin

% Friends: Irene, Sharon, Elayne B
% Sarah Chasins, James Bornholt, Max Willsey, Sami Davies, Daniel Epstein, Elena Agapie, Ravi Karkar, Jessie Schroeder
% Jina 

% Co-authors of this work - Maureen Daum; Nicole, Melissa

% Besties
% Bridget, Aoife, Rachael, Michael

% Partner, Family
% Write Korean - what Mr. Pama said about baksa in our family