\def\adult{\texttt{adult}\xspace}
\def\poundslost{\texttt{pounds\_lost}\xspace}
\def\motivation{\texttt{motivation}\xspace}
\def\regimen{\texttt{regimen\_condition}\xspace}
\def\regimencondition{\texttt{regimen\_condition}\xspace}
\def\group{\texttt{group}\xspace}
\def\age{\texttt{age}\xspace}

Authoring statistical models requires analysts to jointly reason about their
conceptual domain knowledge, statistical methods, and analysis implementations
in code, as our theory of hypothesis formalization describes. For instance,
scientists carefully consider which covariates to include in statistical models
based on their prior knowledge of confounding. However, analysts' conceptual
knowledge is often kept implicit. Analysts gravitate towards statistical
specifications they are familiar with, even if the analyses are sub-optimal or
do not assess their hypotheses, as we saw in the previous chapter. Finally, ease
of implementation further constrains the statistical models that analysts try
and use. These issues are especially salient for domain experts who lack deep
statistical or programming expertise (e.g., many researchers).

Existing statistical software exacerbate these issues because they do not allow
analysts to externalize their implicit conceptual knowledge, receive guidance on
analysis approaches, and help authoring low-level statistical modeling
code~\autoref{sec:toolsAnalysis}. Our work on Tisane hypothesizes that in order
to address these issues, software tools should capture analysts' implicit
conceptual models and use them to derive statistical models. 

\textit{Conceptual models}~\footnote{Richard McElreath calls these implict
assumptions \textit{process models}~\cite{mcelreath2020statistical}. We use the
term \textit{conceptual models} in order to contrast from statistical models.}
are often-informal representations of variable relationships (e.g., list of
variable relationships, process diagrams, graphs), describing the underlying
data generating process. Conceptual models are difficult to reason about during
statistical analysis. Their implications on statistical modeling are not
obvious, especially to statistical non-experts. For example, the impact of
conceptual assumptions may only become apparent after fitting multiple
statistical models, if at all. Without explicitly grappling with conceptual
models prior to authoring statistical models, analysts run the risk of
introducing inconsistencies between their domain knowledge and statistical
models, which can lead to unintentionally answering a different research
question and asserting a conceptual model based on preferred results (i.e.,
HARKing~\cite{}). 

To facilitate more accurate hypothesis formalization and analysis, we asked,
\textbf{How might we derive (initial) statistical models from conceptual
models?}. Inferring a statistical model raises two technical challenges: (1) How
do we elicit the information necessary for inferring a statistical model? and
(2) How do we infer a statistical model, given this information? We explore and
address these issues by iteratively designing, developing, and evaluating
\textbf{Tisane, a system for implementing generalized linear models (GLMs) and
generalized linear mixed-effects models (GLMMs) from explicit statements of
implicit conceptual assumptions}. 

The first implementation of Tisane (\autoref{sec:tisane}) was as an open-source
Python package available on \texttt{pip} (see~\cite{tisaneWebsite}). Case studies and
real-world use of Tisane demonstrated not only the viability but also
the desirability of tool support for authoring statistical models from
conceptual models (\autoref{sec:tisane_case_studies}). Therefore, we explored how to further improve Tisane's
programming and interaction models to better suit novice analysts
(\autoref{sec:exploratoryStudy}) and released a second version in R as the
\rTisane library~\cite{jun2023rTisane} (see~\cite{tisaneWebsite}). The R implementation allowed us to more
directly compare \rTisane to a scaffolded workflow using widely used linear
modeling libraries, including the \texttt{lme4}, in R. 

Tisane provides a \textbf{\SDSLlong} for expressing % conceptual and data
relationships between variables. Tisane compiles the explicitly stated
relationships into an internal \textbf{graph representation} and then traverses
the graph to infer candidate GLMs/GLMMs based on recommendations from the
graphical causal reasoning community. Analysts can then query Tisane for a
statistical model that explains a specific dependent variable from a independent
variable of interest. Tisane helps analysts disambiguate their input conceptual
models and an output statistical model script for fitting a valid GLM/GLMM. In
this way, Tisane focuses analysts on reflecting on and externalizing their
implicit conceptual assumptions and checks that analysts do not overlook
relevant variables, such as potential confounders or data clustering, that could
compromise generalizability of statistical results. %(\higherLevel, \connectConceptualStats). 
% For example, a public health researcher can
% express that average county income is associated with hospital spending based on
% health economics theory or specify that hospitals exist within counties.
%By making conceptual and data measurement knowledge explicit in programs, Tisane protects analysts against specifying statistical models that are inconsistent with domain knowledge and may prompt analysts to reflect on their domain assumptions. % requires analysts to externalize their  domain assumptions, , and avoids statistical model drift from
% The system also guides analysts to examine their
% data distributions, which may challenge and prompt reconsideration of analysts'
% assumptions about a dataset. 
% Figure~\ref{fig:figureSystemOverview} provides an overview
% of this process.

\input{tisane/related_work.tex}
% \section{Tisane's DSL: Formalizing implicit domain assumptions as conceptual models}
% \section{Deriving statistical models from conceptual models}
\input{tisane/system.tex}

\input{tisane/case_studies.tex}

\input{tisane/exploratory.tex}

\input{tisane/rTisane.tex}

\input{tisane/lab_study.tex}

\input{tisane/discussion.tex}

\section{Summary of Contributions}

% Reminder: Thesis statement
% Domain-specific languages that provide abstractions for expressing conceptual
% knowledge, data collection procedures, and analysis intents instead of specific
% statistical modeling decisions coupled with automated reasoning to compile
% conceptual specifications into statistical analysis code help statistical
% non-experts more readily author valid analyses. 

Tisane embodies the hypothesis central to this dissertation: tool support for
expressing implicit conceptual knowledge and reasoning about it to author
statistical analyses enables statistical non-experts to specify valid
statistical models. Through an iterative design process, we refined what the
programming and interaction model for expressing conceptual models and
connecting them to statistical models should be. Most notably, the second
release of \tisane, as \rTisane, provides more explicit support for conceptual
model specification and disambiguation. 

Tisane is a stark contrast to the current ecosystem of statistical analysis
software designed to give analysts maximal mathematical and computational
control at the cost of support for relating their statistical analyses with
their conceptual meaning. The pending lab study results will demonstrate the
impact of \rTisane on (i) the conceptual models analysts specify and their
reflection process, (ii) (output) statistical model quality, and (iii) awareness
and learned insights analysts takeaway about their domain and data analysis
process. We anticipate that our lab study will give evidence for how connecting
conceptual modeling to statistical modeling increases the statistical conclusion
and external validity of analyses~\cite{shadish2010campbell}.  

% The summative evaluation study concretizes the impact of \tisane. Analysts
% report being more reflective and systematic in their thinking about implicit
% conceptual assumptions due to \rTisane's DSL and conceptual model disambiguation
% process. The statistical models analysts produce also more accurately estimate
% the true relationships in the data, lower AIC/BIC and higher R-squared. 


% Prior publications
\textit{The first release of Tisane was a collaboration with Audrey Seo, Jeffrey
Heer, and \reneJust. The corresponding paper was originally published and presented at
\chiConf{2022}~cite{}, where it received a \textit{Best Paper Honorable
Mention Award}. The exploratory design study, second system iteration, and the
summative evaluation are in collaboration with Edward Misback, Jeffrey Heer, and
\reneJust. The second paper is under submission and has not yet been published.}