The previous chapter's case studies (\autoref{sec:tisane_case_studies})
highlighted \tisane's potential as a tool for statistical non-experts to author
generalized linear models with or without mixed effects. To build on \tisane's
strengths, we sought to understand the nuances of how statistical novices wish
to articulate their implicit domain knowledge and the challenges they confront
along the way. We started with a lab study using \tisane to elicit statistical
non-experts' implicit definitions and assumptions about \SDSL keywords
(\autoref{sec:exploratoryStudy}). The study also helped us identify
opportunities to refine \tisane's interactivity. Based on study findings, we
designed and evaluated \textbf{\rTisane, a system to assist novices in
formalizing their conceptual knowledge to author statistical models}. We
implemented \rTisane as an open-source R library based on suggestions from
participants and external research collaborators over the course of the previous
projects. Analysts suggested that tools like \tisane could benefit a wider
audience of novice and expert data analysts if there was an R implementation.
% \rTisane's implementation in R allowed us to reach a potentially wider audience of novice and expert data analysts using R for linear modeling.
An additional advantage of the R implementation was that it allowed us to directly
compare \rTisane to a scaffolded workflow using widely used linear modeling
libraries, including \texttt{lme4}, in R (\autoref{sec:summativeEval}). 


\input{tisane/exploratory.tex}

\input{tisane/rTisane.tex}

\input{tisane/lab_study.tex}

\input{tisane/discussion.tex}

\section{Summary of Contributions}

% Reminder: Thesis statement
% Domain-specific languages that provide abstractions for expressing conceptual
% knowledge, data collection procedures, and analysis intents instead of specific
% statistical modeling decisions coupled with automated reasoning to compile
% conceptual specifications into statistical analysis code help statistical
% non-experts more readily author valid analyses. 

\rTisane provides a DSL with language constructs for expressing conceptual
models (\thesisChallengeExplicit) and integrates a two-phase interactive
disambiguation process for compiling conceptual knowledge into statistical
analysis code (\thesisChallengeRep). In a controlled lab study of \rTisane, we
found that the DSL is expressive enough to capture analysts' conceptual models
accurately, eases the burden of making their implicit assumptions explicit, and
pushes analysts to think about their domains more deeply. Using \rTisane,
analysts were able to author statistical models that fit the data just as well
as if not better than statistical models authored on their own. \rTisane even
helped analysts who were not able to author statistical models on their own get
to an output statistical model. Analysts also reported that through the process,
they learned about GLMs (\thesisChallengeUnderstanding). Together, these results
demonstrate how DSLs and automated reasoning together in fact do help
statistical non-experts author valid statistical analyses that they would not be
able to author otherwise.
% evidence for how
% connecting conceptual modeling to statistical modeling increases the statistical
% conclusion and external validity of analyses~\cite{shadish2010campbell}.  

% The conceptual model disambiguation process in \rTisane
% also facilitates reflection on implicit knowledge. 

% we refined what the programming and interaction model
% for expressing conceptual models and connecting them to statistical models
% should be. Most notably, the second release of \tisane, as \rTisane, provides
% more explicit support for conceptual model specification and disambiguation. 

% \tisane and \rTisane are in stark contrast to the current ecosystem of
% statistical analysis software designed to give analysts maximal mathematical and
% computational control at the cost of support for connecting conceptual and
% statistical models. 
% The pending lab study results will demonstrate the impact of
% \rTisane on (i) the conceptual models analysts specify and their reflection
% process, (ii) (output) statistical model quality, and (iii) awareness and
% learned insights analysts takeaway about their domain and data analysis process.

% The summative evaluation study concretizes the impact of \tisane. Analysts
% report being more reflective and systematic in their thinking about implicit
% conceptual assumptions due to \rTisanes DSL and conceptual model disambiguation
% process. The statistical models analysts produce also more accurately estimate
% the true relationships in the data, lower AIC/BIC and higher R-squared. 


% Prior publications
\textit{The exploratory study, rTisane design and implementation, and the
summative evaluation are in collaboration with Edward Misback, Jeffrey Heer, and
\reneJust. The corresponding paper~\cite{jun2023rTisane} is under submission and has not yet been
published.}
